---
title: "kiki/bouba Bayesian model"
author: "Bodo Winter"
date: "8/8/2019"
output: html_document
---

## Introduction

Load packages:

```{r, message = FALSE}
library(tidyverse)
library(brms)
library(gridExtra)
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

Load the model:

```{r}
load('../models/kiki_mdl.RData')
load('../models/kiki_script_mdl.RData')
load('../models/kiki_field_mdl.RData')
```

Load language data for plotting:

```{r, message = FALSE}
lang_info <- read_csv('../data/language_info.csv')
field_info <- read_csv('../data/language_info_field.csv')
```

## Posterior predictive checks:

Posterior predictive checks:

```{r cache = TRUE, fig.width = 8, fig.height = 6}
pp_check(kiki_mdl, nsample = 1000)
pp_check(kiki_script_mdl, nsample = 1000)
pp_check(kiki_field_mdl, nsample = 1000)
```

## Summarize models:

First, overall summary:

```{r}
summary(kiki_mdl)
```

Extract the posteriors:

```{r}
web_posts <- posterior_samples(kiki_mdl)
```

Get the 95% credible interval in terms of predicted probabilities:

```{r}
plogis(mean(web_posts$b_Intercept))
plogis(quantile(web_posts$b_Intercept, 0.025))
plogis(quantile(web_posts$b_Intercept, 0.975))
```

Posterior predictive probability of the intercept being above chance (logit = 0).

```{r}
sum(web_posts$b_Intercept < 0) / nrow(web_posts)
```

Second, field model:

```{r}
summary(kiki_field_mdl)
```

Extract the posteriors:

```{r}
field_posts <- posterior_samples(kiki_field_mdl)
```

Get the 95% credible interval in terms of predicted probabilities:

```{r}
plogis(mean(field_posts$b_Intercept))
plogis(quantile(field_posts$b_Intercept, 0.025))
plogis(quantile(field_posts$b_Intercept, 0.975))
```

Posterior predictive probability of the intercept being above chance (logit = 0).

```{r}
sum(field_posts$b_Intercept < 0) / nrow(field_posts)
```

## Check orthography-related effects in web-based experiments:

Check script effect in web experiment:

```{r}
summary(kiki_script_mdl)
```

Get the posterior predictions:

```{r}
script_posts <- posterior_samples(kiki_script_mdl)
```

Check whether the coefficients are reliably different from 0:

```{r}
sum(script_posts$b_Scriptroman > 0) / nrow(script_posts)
```

What is the predicted increase in probability for roman script?

```{r}
# Non-roman script:

plogis(fixef(kiki_script_mdl)[1, 1])

# Roman script:

plogis(fixef(kiki_script_mdl)[1, 1] + fixef(kiki_script_mdl)[2, 1])

# Odds:

exp(fixef(kiki_script_mdl)[2, 1])
```

## Plot of predicted ACC by language, web experiment:

Get the posteriors for the average effects:

```{r}
lang_samples <- select(web_posts, contains('Language'))
lang_samples <- lang_samples[, -1]
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(lang_samples, FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(lang_samples, FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(lang_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
lang_intervals <- tibble(Language = str_extract(colnames(lang_samples), '[A-Z]{2,2}'),
                         lower_CI = plogis(lower_CIs),
                         upper_CI = plogis(upper_CIs),
                         M = plogis(post_mean))
```

Merge with name for plotting:

```{r}
lang_intervals$Name <- lang_info[match(lang_intervals$Language, lang_info$Language), ]$Name
```

Change "English UK" label to just "English":

```{r}
lang_intervals[lang_intervals$Name == 'English UK', ]$Name <- 'English'
```

Change the order of language levels according to the posterior mean:

```{r}
lang_intervals <- arrange(lang_intervals, M)
lang_intervals <- mutate(lang_intervals,
                         Name = factor(Name, levels = lang_intervals$Name))
```

Make a plot with accuracy per language:

```{r, fig.width = 10, fig.height = 6}
lang_plot <- ggplot(data = lang_intervals,
                    mapping = aes(x = Name, y = M, ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'goldenrod3', col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/2, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 24.5, y = 1/2 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(a) Estimated accuracy by language, web experiment',
       x = '', y = 'Posterior probability')
lang_plot
```

## Plot of ACC per language, field data:

Getting the posteriors per language:

```{r}
field_lang_samples <- select(field_posts, contains('Language'))
head(field_lang_samples)
```

Get rid of the first, which is across all of them:

```{r}
field_lang_samples <- field_lang_samples[, -1]
```

Compute 95% CI for all of these:

```{r}
lower_CIs <- apply(field_lang_samples, FUN = function(x) quantile(x, 0.025), MARGIN = 2)
upper_CIs <- apply(field_lang_samples, FUN = function(x) quantile(x, 0.975), MARGIN = 2)
post_mean <- colMeans(field_lang_samples)
```

Put those into a table, with logistic transform back to probabilities:

```{r}
field_lang_intervals <- tibble(Language = str_extract(colnames(field_lang_samples), '[A-Z]{2,2}'),
                               lower_CI = plogis(lower_CIs),
                               upper_CI = plogis(upper_CIs),
                               M = plogis(post_mean))
```

Add language names for plotting:

```{r}
field_lang_intervals$Name <- NA
field_lang_intervals[field_lang_intervals$Language == 'DE', ]$Name <- 'German'
field_lang_intervals[field_lang_intervals$Language == 'BE', ]$Name <- 'English (UK)'
field_lang_intervals[field_lang_intervals$Language == 'BR', ]$Name <- 'Tashlhiyt Berber'
field_lang_intervals[field_lang_intervals$Language == 'PA', ]$Name <- 'Palikur'
field_lang_intervals[field_lang_intervals$Language == 'VA', ]$Name <- 'Daakie'
field_lang_intervals[field_lang_intervals$Language == 'SR', ]$Name <- 'Brazilian Portuguese'
```

Change the order of language levels according to the posterior mean:

```{r}
field_lang_intervals <- arrange(field_lang_intervals, M)
field_lang_intervals <- mutate(field_lang_intervals,
                         Name = factor(Name, levels = field_lang_intervals$Name))
```

Make a plot with accuracy per language:

```{r, fig.width = 10, fig.height = 6}
field_lang_plot <- ggplot(data = field_lang_intervals,
                    mapping = aes(x = Name, y = M, ymin = lower_CI, ymax = upper_CI)) +
  geom_errorbar(width = 0.3, lwd = 1) +
  geom_point(pch = 22, size = 7, fill = 'goldenrod3', col = 'black', stroke = 1) +
  geom_hline(yintercept = 1/2, linetype = 2) +
  theme_timo +
  annotate('text', label = 'Chance level', x = 7.5, y = 1/2 + 0.04,
           size = 10, fontface = 'bold', hjust = 1) +
  coord_cartesian(ylim = c(0, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = 'bold', size = 24),
        axis.title.y = element_text(margin = margin(t = 0, r = 50, b = 0, l = 0),
                                    size = 30),
        axis.text.y = element_text(face = 'bold', size = 20),
        plot.title = element_text(face = 'bold', size = 36, hjust = 0,
                                  margin = margin(t = 0, r = 0, b = 20, l = 0))) +
  labs(title = '(b) Estimated accuracy by language, field experiment',
       x = '', y = 'Posterior probability')
field_lang_plot
```

Create a plot of both of these:

```{r}
my_layout <- rbind(c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(NA, NA, NA, NA, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA))

field_multi_plot <- grid.arrange(lang_plot, field_lang_plot,
                           layout_matrix = my_layout)
ggsave(plot = field_multi_plot, '../figures/posterior_means_by_language.png', width = 20, height = 22)
ggsave(plot = field_multi_plot, '../figures/posterior_means_by_language.pdf', width = 20, height = 22)
```

This completes this analysis.

