---
title: "kiki/bouba Bayesian model"
author: "Bodo Winter"
date: "8/8/2019"
output: html_document
---

## Introduction

This script takes the models that are produced by "bayes_models.R" and reports summary results:

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(brms)
library(gridExtra)
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

Get data to plot descriptive averages to the model:

```{r, message = FALSE, warning = FALSE}
kiki <- read_csv('../data/web_experiment_cleaned.csv')
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

Load the model:

```{r}
load('../models/kiki_mdl.RData')
load('../models/kiki_script_mdl.RData')
load('../models/daakie_mdl.RData')
load('../models/kiki_AnyRoman_mdl.RData')
load('../models/kiki_scriptL2_mdl.RData')
```

Load language data for plotting:

```{r, message = FALSE}
lang_info <- read_csv('../data/language_info.csv')
```

## Posterior predictive checks:

Posterior predictive checks:

```{r cache = TRUE, fig.width = 8, fig.height = 6}
pp_check(kiki_mdl, nsample = 1000)
pp_check(kiki_script_mdl, nsample = 1000)
pp_check(kiki_AnyRoman_mdl, nsample = 1000)
pp_check(kiki_scriptL2_mdl, nsample = 1000)
pp_check(daakie_mdl, nsample = 1000)
```

## Summarize models:

First, overall summary:

```{r}
summary(kiki_mdl)
```

Extract the posteriors:

```{r}
web_posts <- posterior_samples(kiki_mdl)
```

Get the 95% credible interval in terms of predicted probabilities:

```{r}
plogis(mean(web_posts$b_Intercept))
plogis(quantile(web_posts$b_Intercept, 0.025))
plogis(quantile(web_posts$b_Intercept, 0.975))
```

Posterior predictive probability of the intercept being above chance (logit = 0).

```{r}
sum(web_posts$b_Intercept < 0) / nrow(web_posts)
```

## Check orthography-related effects in web-based experiments:

Check script effect in web experiment:

```{r}
summary(kiki_script_mdl)
```

Get the posterior predictions:

```{r}
script_posts <- posterior_samples(kiki_script_mdl)
```

Check whether the coefficients are reliably different from 0:

```{r}
sum(script_posts$b_Scriptroman < 0) / nrow(script_posts)
```

What is the predicted increase in probability for roman script?

```{r}
# Non-roman script:

plogis(fixef(kiki_script_mdl)[1, 1])

# Roman script:

plogis(fixef(kiki_script_mdl)[1, 1] + fixef(kiki_script_mdl)[2, 1])

# Odds:

exp(fixef(kiki_script_mdl)[2, 1])
```


## Check the L2 related effects:

Check script effect in web experiment:

```{r}
summary(kiki_AnyRoman_mdl)
```

Get the posterior predictions:

```{r}
L2_posts <- posterior_samples(kiki_AnyRoman_mdl)
```

Intercept?

```{r}
sum(L2_posts$b_Intercept > 0) / nrow(L2_posts)
```

## Check script L2 effect:

Summarize model:

```{r}
summary(kiki_scriptL2_mdl)
```

Get the posterior predictions:

```{r}
scriptL2_posts <- posterior_samples(kiki_scriptL2_mdl)
```

Check whether the coefficients are reliably different from 0:

```{r}
sum(scriptL2_posts$b_factorRomanScriptL2TRUE < 0) / nrow(scriptL2_posts)
```

What is the predicted increase in probability for roman script?

```{r}
# Non-roman script:

plogis(fixef(kiki_scriptL2_mdl)[1, 1])

# Roman script:

plogis(fixef(kiki_scriptL2_mdl)[1, 1] + fixef(kiki_scriptL2_mdl)[2, 1])

# Odds:

exp(fixef(kiki_scriptL2_mdl)[2, 1])
```



## Check Daakie speakers:

Check model:

```{r}
summary(daakie_mdl)
```

Check 95% CI:

```{r}
plogis(fixef(daakie_mdl)[3])
plogis(fixef(daakie_mdl)[4])
```

Check posterior probability of the effect being below 0:

```{r}
daakie_posts <- posterior_samples(daakie_mdl)
sum(daakie_posts$b_Intercept < 0) / nrow(daakie_posts)
```


## Make a plot of the main model predictions:

Extract language random effects:

```{r}
language_posts <- select(web_posts, contains('r_Language'))
colnames(language_posts) <- str_extract(colnames(language_posts), '[A-Z]{2,2}')
```

Add intercept to all:

```{r}
language_posts <- language_posts + web_posts$b_Intercept
```

Extract posterior means and 95% credible interval:

```{r}
preds <- tibble(Lang_ID = colnames(language_posts),
                Mean = plogis(colMeans(language_posts)))
```

Name the languages more transparently:

```{r}
preds$Language <- kiki[match(preds$Lang_ID, kiki$Language), ]$Name
```

Get 95% CIs:

```{r}
preds$Lower_CI <- apply(language_posts, 2, FUN = function(x) quantile(x, 0.025))
preds$Upper_CI <- apply(language_posts, 2, FUN = function(x) quantile(x, 0.975))
```

Transform them into proportions:

```{r}
preds <- mutate(preds, Lower_CI = plogis(Lower_CI),
                Upper_CI = plogis(Upper_CI))
```

Add script info:

```{r}
preds$Script <- kiki[match(preds$Lang_ID, kiki$Language), ]$Script
preds <- mutate(preds,
                Script = ifelse(Script == 'roman', 'Roman script', 'other script'))
```

Add descriptive averages:

```{r}
kiki_avgs <- kiki %>% group_by(Language) %>%
  summarize(ACC = mean(Match))

# Merge:

preds$ACC <- kiki_avgs$ACC
```

Sort the languages by posterior means, and factor-code them in that order too:

```{r}
preds <- arrange(preds, Mean) %>%
  mutate(Language = fct_reorder(Language, Mean))
```

Sort script factor so it's displayed in the right order:

```{r}
preds <- mutate(preds,
                Script = factor(Script, levels = c('Roman script', 'other script')))
```


Make a plot of this:

```{r, fig.width = 14, fig.height = 5}
# Aesthetics:

web_p <- preds %>%
  ggplot(aes(x = reorder(Language, Mean),
             y = Mean, col = Script))

# Geoms:

web_p <- web_p +
  geom_hline(yintercept = 0.5, linetype = 2, size = 1.5, col = 'grey') + 
  geom_point(size = 6, shape = 15) +
  geom_errorbar(mapping = aes(ymin = Lower_CI, ymax = Upper_CI), size = 1.4) +
  geom_point(aes(y = ACC), col = 'black', shape = 18, size = 6)
  
# Extra stuff:

web_p <- web_p + 
  labs(x = '', y = 'Proportion\n') +
  ggtitle('(c) Posterior means and descriptive averages of matching responses by language and script') + 
  coord_cartesian(ylim = c(0, 1)) + 
  scale_color_viridis_d(option = 'D', begin = 0.3, end = 0.8, direction = 1) +
  theme_timo +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = 'bold',
                                   size = 30),
        axis.text.y = element_text(face = 'bold', size = 24),
        axis.title = element_text(face = 'bold', size = 40),
        axis.title.y = element_text(face = 'bold', size = 40),
        plot.title = element_text(face = 'bold', size = 40,
                                  margin = margin(t = 0, r = 0,
                                                  b = 20, l = 0)),
        legend.text = element_text(size = 30),
        legend.title = element_blank(),
        legend.position = c(0.95, 0.15),
        legend.justification = c('right', 'bottom'))

# Show:

web_p

# Save:

ggsave(plot = web_p, filename = '../figures/Figure1c_Bayes_descriptive.pdf',
       width = 28, height = 12)
```

This completes this analysis.

