---
title: "Descriptive stats for kiki/bouba experiments (field + web)"
author: "Bodo Winter"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the preprocessing script that cleans the data from the online and the field work experiment.

## Load in data

Load packages:

```{r, message = FALSE}
library(tidyverse)
library(lsr)    # For calculating Cramer's V 
library(gridExtra)
```

Load data:

```{r, message = FALSE, warning = FALSE}
kiki <- read_csv('../data/web_experiment_cleaned.csv')
field <- read_csv('../data/field_experiment_cleaned.csv')
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

## Overall stats:

How may speakers?

```{r}
# Web experiment:

length(unique(kiki$ID))

# Field experiment:

length(unique(field$ID))
```

How many languages?

```{r}
# Web experiment:

length(unique(kiki$Language))

# Field experiment:

length(unique(field$Language))
```

Total number of languages across experiments?

```{r}
# Minus German, English, Portuguese, which are counted doubly as they occur in both:

length(unique(kiki$Language)) + length(unique(field$Language)) - 3
```

How many language families?

```{r}
# +1 because Palikur from the field data is Arawakan, which is a language family that is not in the web data

length(unique(kiki$Family)) + 1
```

How many Indo-European languages?

```{r}
unique(filter(kiki, Family == 'IE')$Name)
length(unique(filter(kiki, Family == 'IE')$Name))
```

What are the non-Indo-European language families?

```{r}
unique(filter(kiki, Family != 'IE')$Family)
```

How many per language?

```{r}
# Web experiment:

table(filter(kiki, !duplicated(ID))$Language)

# Field experiment:

table(filter(field, !duplicated(ID))$Language)
```

## Web experiment:

Check the overall proportion of matches:

```{r}
round(mean(kiki$Match, na.rm = TRUE), 2)
```


Overall 71% matching the hypothesis (that "kiki" input = kiki shape, and "bouba" input = bouba shape).

Check this for the different languages in separate mini tables:

```{r}
lang_tab <- kiki %>%
  group_by(Language, Name, Family, Script) %>%
  summarize(Match = mean(Match), Match = round(Match, 2))

lang_tab <- kiki %>% count(Language) %>% right_join(lang_tab)

lang_tab <- lang_tab %>% 
  arrange(desc(Match))

lang_tab %>% 
  print(n = Inf)
```

Get the average proportions by script:

```{r}
kiki %>% group_by(Script) %>% 
  summarize(Match = mean(Match))
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
lang_tab %>%
  ggplot(aes(x = reorder(Language, Match),
             y = Match, fill = Script)) +
  geom_col() +
  labs(x = '', y = 'Proportion\n') +
  ggtitle('Proportion match by language') + 
  coord_cartesian(ylim = c(0, 1)) + 
  geom_hline(yintercept = 0.5, linetype = 2, size = 2) + 
  scale_fill_viridis_d(option = 'D', begin = 0.3, end = 0.8) +
  theme_timo
```

This suggests that languages with different scripts on average seem to have less strong of a kiki/bouba effect. Of course, that's partially confounded with genealogical associations since languages with different scripts are more likely non-Indo-European.

Create a variable for IE versus non-IE:

```{r}
lang_tab <- mutate(lang_tab,
                   `Language Family` = ifelse(Family == 'IE', 'IE', 'other'))
```

Change the script label so it's nice for plotting:

```{r}
lang_tab <- mutate(lang_tab,
                   Script = ifelse(Script == 'roman', 'Roman script', 'non-Roman script'))
```

Make a plot broken up by IE versus non-IE:

```{r, fig.width = 12, fig.height = 12}
# Aesthetics:

web_p <- lang_tab %>%
  ggplot(aes(x = reorder(Name, Match),
             y = Match, fill = Script))

# Geoms:

web_p <- web_p +
  geom_bar(stat = 'identity', width = 0.7) +
  geom_hline(yintercept = 0.5, linetype = 2, size = 2)

# Extra stuff:

web_p <- web_p + 
  labs(x = '', y = 'Proportion\n') +
  ggtitle('(a) Web experiment') + 
  coord_cartesian(ylim = c(0, 1)) + 
  scale_fill_viridis_d(option = 'D', begin = 0.3, end = 0.8, direction = 1) +
  theme_timo +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = 'bold', size = 24),
        axis.text.y = element_text(face = 'bold', size = 24),
        axis.title = element_text(face = 'bold', size = 40),
        axis.title.y = element_text(face = 'bold', size = 40),
        plot.title = element_text(face = 'bold', size = 40,
                                  margin = margin(t = 0, r = 0,
                                                  b = 20, l = 0)),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 32, face = 'bold'))

# Show:

web_p
```

## Field experiment:

First, overall match:

```{r}
mean(field$Match, na.rm = TRUE)
```

Check result depending on whether modality was auditory or written (aggregating across languages and speakers):

```{r}
field %>% group_by(Modality) %>% 
  summarize(Match = mean(Match, na.rm = TRUE))
```

Not much going on here. A mere 3% difference.

For various reasons, we want to focus on the spoken version of the task. This is also because it was always run first.

```{r}
field <- filter(field, Modality == 'Aud')
```

Check results by languages:

```{r}
field_lang_tab <- field %>% 
  group_by(Language) %>%
  summarize(Match = mean(Match, na.rm = TRUE),
            Match = round(Match, 2))

field_lang_tab <- field %>% count(Language) %>% right_join(field_lang_tab)

field_lang_tab <- field_lang_tab %>% 
  arrange(desc(Match))

field_lang_tab
```

For plot, create full language names:

```{r}
field_lang_tab$Name <- c('British English',
                         'Tashlhiyt Berber',
                         'German',
                         'Daakie',
                         'Palikur',
                         'Brazilian Portuguese')
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
# Aesthetics:

field_p <- field_lang_tab %>%
  ggplot(aes(x = reorder(Name, Match),
             y = Match))

# Geoms:

field_p <- field_p + 
  geom_bar(fill = 'gray42', stat = 'identity', width = 0.8)

# Extra stuff:

field_p <- field_p +
  labs(x = '', y = 'Proportion\n') +
  ggtitle('(b) Field experiment') +
  coord_cartesian(ylim = c(0, 1)) + 
  geom_hline(yintercept = 0.5, linetype = 2, size = 2) + 
  theme_timo +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = 'bold', size = 24),
        axis.text.y = element_text(face = 'bold', size = 24),
        axis.title = element_text(face = 'bold', size = 40),
        axis.title.y = element_text(face = 'bold', size = 40),
        legend.position = 'none',
        plot.title = element_text(face = 'bold', size = 40,
                                  margin = margin(t = 0, r = 0,
                                                  b = 20, l = 0)))

# Plot:

field_p
```

## Combine both plots:

First, create a layout matrix:

```{r}
my_layout <- rbind(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA))
```

Then, create the double plot:

```{r, fig.width = 14, fig.height = 10}
multi_plot <- grid.arrange(web_p, field_p,
                           layout_matrix = my_layout)
ggsave(plot = multi_plot, '../figures/per_language_barplots.png', width = 24, height = 20)
ggsave(plot = multi_plot, '../figures/per_language_barplots.pdf', width = 24, height = 20)
```

This completes this analysis.