---
title: "Descriptive stats for kiki/bouba experiments (field + web)"
author: "Bodo Winter"
date: "7/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is script performs descriptive statistics of the online data and the field work experiment.

Note on the field experiment:
The field work experiment is not reported in the main analysis. We do find a clear kiki/bouba effect in the field work data, however, there was a critical error in the procedure: the written version was always run before the spoken version. As a result of this, the spoken versus written comparison is not interpretable, as the spoken results could be biased by having seen a prior orthographic representation. Moreover, the web experiment and the field experiment are thus not comparable anymore.

## Load in data

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(lsr)    # For calculating Cramer's V 
library(gridExtra)

# For mapping:

library(sf)
library(ggpubr)
library(rnaturalearth)
library(rnaturalearthdata)
```

Load data:

```{r, message = FALSE, warning = FALSE}
kiki <- read_csv('../data/web_experiment_cleaned.csv')
field <- read_csv('../data/field_experiment_cleaned.csv')
```

Load ggplot2 theme:

```{r}
source('theme_timo.R')
```

## Overall stats:

How may speakers?

```{r}
# Web experiment:

length(unique(kiki$ID))

# Field experiment:

length(unique(field$ID))
```

How many languages?

```{r}
# Web experiment:

length(unique(kiki$Language))

# Field experiment:

length(unique(field$Language))
```

Total number of languages across experiments?

```{r}
# Minus German, English, Portuguese, which are counted doubly as they occur in both:

length(unique(kiki$Language)) + length(unique(field$Language)) - 3
```

How many language families?

```{r}
length(unique(kiki$Family))
```

How many Indo-European languages?

```{r}
unique(filter(kiki, Family == 'IE')$Name)
length(unique(filter(kiki, Family == 'IE')$Name))
```

What are the non-Indo-European language families?

```{r}
unique(filter(kiki, Family != 'IE')$Family)
length(unique(filter(kiki, Family != 'IE')$Name))
```

How many per language?

```{r}
# Web experiment:

table(filter(kiki, !duplicated(ID))$Language)

# Field experiment:

table(filter(field, !duplicated(ID))$Language)
```

How many languages that have non-Roman scripts?

```{r}
unique(kiki[kiki$Script != 'roman', ]$Name)
length(unique(kiki[kiki$Script != 'roman', ]$Language))

# How many of the total is that?

length(unique(kiki[kiki$Script != 'roman', ]$Language)) / length(unique(kiki$Language))
```

9 languages out of 24, more than a third, have non-Roman scripts.

How many participants are bilingual?

```{r}
sum(is.na(kiki$L2))
sum(!is.na(kiki$L2)) # knows an L2

round(sum(is.na(kiki$L2)) / nrow(kiki), 2)
round(sum(!is.na(kiki$L2)) / nrow(kiki), 2)
```

How many of the bilingual ones speak english as L2?

```{r}
kiki %>% filter(Language != 'EN') %>% 
  count(EnglishL2YesNo) %>% mutate(prop = n / sum(n),
                                   prop = round(prop, 2))
```

How many participants know English?

```{r}
kiki %>% count(EnglishL2YesNo) %>%
  mutate(Prop = n / sum(n),
         Prop = round(Prop, 2))
```

How many participants know a language with Roman alphabet as L2?

```{r}
kiki %>% count(RomanScriptL2) %>%
  mutate(Prop = n / sum(n),
         Prop = round(Prop, 2))
```

How many participants know a non-Indo-European language as L2?

```{r}
kiki %>% count(IEL2) %>%
  mutate(Prop = n / sum(n),
         Prop = round(Prop, 2))
```

How many speakers are "pure non-IE speakers", that is, they speak a non-IE language as first language and no IE as second language?

```{r}
kiki %>% filter(Family != 'IE') %>% count(IEL2)
```

How many speakers are "pure non-Roman script writers", that is, they speak no language that contains the Roman script??

```{r}
kiki %>% filter(Script != 'roman') %>% count(RomanScriptL2)
```

## Web experiment:

Check the overall proportion of matches:

```{r}
round(mean(kiki$Match, na.rm = TRUE), 2)
```

Overall 71% matching the hypothesis (that "kiki" input = kiki shape, and "bouba" input = bouba shape).

Check this for the different L2:

```{r}
kiki %>%
  filter(Language != 'EN') %>%
  group_by(EnglishL2YesNo) %>% 
  summarize(Match = round(mean(Match, na.rm = TRUE), 2))
```

Check this for whether the L2 spoken has a Roman alphabet:

```{r}
kiki %>%
  group_by(RomanScriptL2) %>% 
  summarize(Match = round(mean(Match, na.rm = TRUE), 2))
```

Check this for whether the L2 spoken is non-IE:

```{r}
kiki %>%
  filter(Language != 'EN') %>%
  group_by(IEL2) %>% 
  summarize(Match = round(mean(Match, na.rm = TRUE), 2))
```

All of these are much above chance as well.

Let's check the "pure" speakers though:

```{r}
kiki %>% filter(Family != 'IE') %>% filter(!IEL2) %>% 
  summarize(Match = round(mean(Match, na.rm = TRUE), 2))
```

65%... still much above 50%.

How many speakers are "pure non-Roman script writers", that is, they speak no language that contains the Roman script??

```{r}
kiki %>% filter(Script != 'roman') %>% filter(!RomanScriptL2) %>% 
  summarize(Match = round(mean(Match, na.rm = TRUE), 2))
```

Still 67%, which is much above 50%.

```{r}
binom.test(c(19, 38))
```

This would be reliable per binomial test..

Check this for the different languages in separate mini tables:

```{r}
lang_tab <- kiki %>%
  group_by(Language, Name, Family, Script) %>%
  summarize(Match = mean(Match), Match = round(Match, 2))

lang_tab <- kiki %>% count(Language) %>% right_join(lang_tab)

lang_tab <- lang_tab %>% 
  arrange(desc(Match))

lang_tab %>% 
  print(n = Inf)
```

Get the average proportions by script:

```{r}
kiki %>% group_by(Script) %>% 
  summarize(Match = mean(Match))
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
lang_tab %>%
  ggplot(aes(x = reorder(Language, Match),
             y = Match, fill = Script)) +
  geom_col() +
  labs(x = '', y = 'Proportion\n') +
  ggtitle('Proportion match by language') + 
  coord_cartesian(ylim = c(0, 1)) + 
  geom_hline(yintercept = 0.5, linetype = 2, size = 2) + 
  scale_fill_viridis_d(option = 'D', begin = 0.3, end = 0.8) +
  theme_timo
```

This suggests that languages with different scripts on average seem to have less strong of a kiki/bouba effect. Of course, that's partially confounded with genealogical associations since languages with different scripts are more likely non-Indo-European.

Create a variable for IE versus non-IE:

```{r}
lang_tab <- mutate(lang_tab,
                   `Language Family` = ifelse(Family == 'IE', 'IE', 'other'))
```

Change the script label so it's nice for plotting:

```{r}
lang_tab <- mutate(lang_tab,
                   Script = ifelse(Script == 'roman', 'Roman script', 'non-Roman script'))
```

Make a plot broken up by IE versus non-IE:

```{r, fig.width = 12, fig.height = 12}
# Aesthetics:

web_p <- lang_tab %>%
  ggplot(aes(x = reorder(Name, Match),
             y = Match, fill = Script))

# Geoms:

web_p <- web_p +
  geom_bar(stat = 'identity', width = 0.7) +
  geom_hline(yintercept = 0.5, linetype = 2, size = 2)

# Extra stuff:

web_p <- web_p + 
  labs(x = '', y = 'Proportion\n') +
  ggtitle('(a) Web experiment') + 
  coord_cartesian(ylim = c(0, 1)) + 
  scale_fill_viridis_d(option = 'D', begin = 0.3, end = 0.8, direction = 1) +
  theme_timo +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = 'bold', size = 24),
        axis.text.y = element_text(face = 'bold', size = 24),
        axis.title = element_text(face = 'bold', size = 40),
        axis.title.y = element_text(face = 'bold', size = 40),
        plot.title = element_text(face = 'bold', size = 40,
                                  margin = margin(t = 0, r = 0,
                                                  b = 20, l = 0)),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 32, face = 'bold'))

# Show:

web_p
```

## Field experiment:

For the field experiment, we had the planning issue that in all cases except Daakie, the written experiment was run first. This makes this data not open to testing the idea that there is a genuine effect of spoken iconicity, as the spoken responses are contaminated because of having seen the spiky and round letters first.

Because of this, we are only going to look at Daakie for the analysis reported in the paper.

It should be noted that the results from the other languages do support the kiki/bouba effect, but only the ritten version 

Let's get only the Daakie data points:

```{r}
field_red <- filter(field, Language == 'VA', Modality == 'Aud')
```

First, overall match:

```{r}
str_c(round(mean(field_red$Match), 3) * 100, '%')
```

This is such little data, let's make a table:

```{r}
table(field_red$Match)
```

Write this to table for Bayesian analysis:

```{r}
write_csv(field_red, '../data/field_daakie_only.csv')
```

## Analysis of the remaining data (not reported in paper):

First, overall match:

```{r}
mean(field$Match, na.rm = TRUE)
```

Check result depending on whether modality was auditory or written (aggregating across languages and speakers):

```{r}
field %>% group_by(Modality) %>% 
  summarize(Match = mean(Match, na.rm = TRUE))
```

Not much going on here. A mere 3% difference.

For various reasons, we want to focus on the spoken version of the task. This is also because it was always run first.

```{r}
field <- filter(field, Modality == 'Aud')
```

Check results by languages:

```{r}
field_lang_tab <- field %>% 
  group_by(Language) %>%
  summarize(Match = mean(Match, na.rm = TRUE),
            Match = round(Match, 2))

field_lang_tab <- field %>% count(Language) %>% right_join(field_lang_tab)

field_lang_tab <- field_lang_tab %>% 
  arrange(desc(Match))

field_lang_tab
```

For plot, create full language names:

```{r}
field_lang_tab$Name <- c('British English',
                         'Tashlhiyt Berber',
                         'German',
                         'Daakie',
                         'Palikur',
                         'Brazilian Portuguese')
```

Make a plot of this:

```{r, fig.width = 12, fig.height = 6}
# Aesthetics:

field_p <- field_lang_tab %>%
  ggplot(aes(x = reorder(Name, Match),
             y = Match))

# Geoms:

field_p <- field_p + 
  geom_bar(fill = 'gray42', stat = 'identity', width = 0.8)

# Extra stuff:

field_p <- field_p +
  labs(x = '', y = 'Proportion\n') +
  ggtitle('(b) Field experiment') +
  coord_cartesian(ylim = c(0, 1)) + 
  geom_hline(yintercept = 0.5, linetype = 2, size = 2) + 
  theme_timo +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = 'bold', size = 24),
        axis.text.y = element_text(face = 'bold', size = 24),
        axis.title = element_text(face = 'bold', size = 40),
        axis.title.y = element_text(face = 'bold', size = 40),
        legend.position = 'none',
        plot.title = element_text(face = 'bold', size = 40,
                                  margin = margin(t = 0, r = 0,
                                                  b = 20, l = 0)))

# Plot:

field_p
```

## Combine both plots:

First, create a layout matrix:

```{r}
my_layout <- rbind(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                   c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA),
                   c(2, 2, 2, 2, 2, 2, NA, NA, NA, NA, NA))
```

Then, create the double plot:

```{r, fig.width = 14, fig.height = 10}
multi_plot <- grid.arrange(web_p, field_p,
                           layout_matrix = my_layout)
ggsave(plot = multi_plot, '../figures/per_language_barplots.png', width = 24, height = 20)
ggsave(plot = multi_plot, '../figures/per_language_barplots.pdf', width = 24, height = 20)
```

This completes this analysis.

## Plot of the world map:

Load country-level data:

```{r, message = FALSE, warning = FALSE}
web_countries <- read_csv('../data/countries_map_online.csv')

# From: https://developers.google.com/public-data/docs/canonical/countries_csv

canonical <- read_csv('../data/countries.csv')
```

Merge the longitude and latitude info in there:

```{r}
web_countries <- left_join(web_countries, canonical,
                           by = c('CountryCode' = 'country'))
```

Make a map:

```{r, fig.width = 9, fig.height = 6}
world <- ne_countries(scale = "medium", returnclass = "sf")

world_p <- ggplot(world) +
  geom_sf(size=0.2) +
  coord_sf(xlim = c(-180, 180), ylim = c(-60, 80), datum=NA) +
  geom_point(data = web_countries,
             aes(x = longitude, y = latitude, group = NA),
             alpha = 0.8, position = position_jitter(0.1), size = 3,
             col = 'firebrick1') +
  labs(title = "(b) Countries from the participants of our web survey") +
  theme_timo +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0.1,0.4,-0.4,0.4),"cm"),
    panel.spacing = unit(0, "lines"),
    plot.title = element_text(size = 12))

# Plot:

world_p

# Save:

ggsave('../figures/Figure2_world_map.pdf', plot = world_p)
```



